Software Citation Principles

### 2016-03-31 sc-principles/software-citation-principles.tex
fix typo
edits to discussion of software papers based on #84

While not usually expressed as software citation priciples, a number of groups have developed community guidelines around software and data citation. Van de Sompel et al.~\cite{VandeSompel2004} argue for registration of all units of scholarly communication, including software. In a `Publish or be damned? An alternative impact manifesto for research software'~\cite{ssi-publish-or-be-damned}, Chue Hong lists nine principles as part of `The Research Software Impact Manifesto'.

@@

Currently, and for the foreseeable future, software papers are being published and cited,
in addition to software itself being cited, as many community norms and practices are
oriented towards citation of papers.
Given a software paper and software, the question of which should be cited and when is
complicated, and there is no simple answer.
As discussed in the Importance principle (\ref{principle:importance}), \textit{the software
itself should also be cited if it is directly used and was important to the work}.
If a software paper exists and it contains results (performance, validation, etc.) that are
important to the work, then the software paper should also be cited.
In addition, if the software authors ask that a paper should be cited, that should be respected.
Note that there are situations where it is appropriate to cite the software, the software paper, or both.

>>>>

While not usually expressed as software citation principles, a number of groups have developed community guidelines around software and data citation. Van de Sompel et al.~\cite{VandeSompel2004} argue for registration of all units of scholarly communication, including software. In a `Publish or be damned? An alternative impact manifesto for research software'~\cite{ssi-publish-or-be-damned}, Chue Hong lists nine principles as part of `The Research Software Impact Manifesto'.

@@

Currently, and for the foreseeable future, software papers are being published and cited,
in addition to software itself being cited, as many community norms and practices are
oriented towards citation of papers.
Where software should be cited, the question of whether to cite the software itself or a software paper is
complicated, and there is no simple answer.
As discussed in the Importance principle (\ref{principle:importance}), \textit{the software
itself should also be cited if it is directly used and was important to the work}.
If a software paper exists and it contains results (performance, validation, etc.) that are
important to the work, then the software paper should also be cited.
In addition, if the software authors ask that a paper should be cited, that should typically be respected and the paper cited in addition to the software. 
However, the software itself should still be cited.
Note that there are situations where it is appropriate to cite the software, the software paper, or both.
